### 2014-11-15 Articles_of_Association_Melbourne_Bitcoin_Technology_Centre_MBTC_2015.tex
Draft first page preamble and definitions

% --------------
% LaTeX Document
% --------------

\documentclass[english,a4paper,titlepage,oneside]{article}

% --------
% PREAMBLE
% --------

% Use Packages
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{fontenc}
\usepackage[top=2.25cm, bottom=3cm, left=2.5cm, right=2.5cm]{geometry}

% Set paragraph vert space and indetatoin.
\setlength\parskip{\medskipamount}
\setlength\parindent{0pt}

% Assert paragraph unit formatting
\raggedright

% Hyphen dissuasion threshold
\hyphenpenalty=10000

% Define interperatation of @ charachter
\makeatletter
\makeatother

% ----------------
% DOCUMENT CONTENT
% ----------------

\begin{document}

% Letter
\section*{\centerline{Articles of Association of}}
\section*{\centerline{BITCOIN TECHNOLOGY CENTERS, INC.}}
\section*{\centerline{Incorporated in Victoria under the}}
\section*{\centerline{Associations Incorporation Reform Act 2012}}

The regulations as were in force upon the date of adoption of these Articles shall apply to the Association save to the extent that they are excluded by or are inconsistent with any of these Articles.

\section*{Interpretation}

\begin{enumerate}
\item In these articles:

\begin{itemize}
\item {}``\emph{the Act}'' means the Associations Incorporation Reform Act 
2012 including the Associations Incorporation Reform Regulations 2012;
\item {}``\emph{address}'' means a postal address or, for the purposes
of electronic communication, an e-mail address or a addressable 
communication channel in each case registered with or by the Association;
\item {}``\emph{the Association}'' means the association intended to be bound by these articles;
\item {}``\emph{clear days}'' in relation to the period of a notice means
a period excluding:

\begin{itemize}
\item the day when the notice is given or deemed to be given; and
\item the day for which it is given or on which it is to take effect;
\end{itemize}
\item {}``\emph{the Articles}'' means the Articles of Association
of the Association;
\item {}``\emph{officers}'' includes the Board and the Secretary;
\item {}``\emph{the seal}'' means the common seal of the Association if it
has one;
\item {}``\emph{Secretary}'' means the Secretary of the Association or any
other person appointed to perform the duties of the Secretary of the
Association, including a joint, assistant or deputy secretary;
\item {}``\emph{the Board}'' means the members of the Association so appointed. During any period in which the Association is a charity, the Board members are trustees as defined by Section the Charities Act 1997;
\item {}``\emph{Victoria}'' means the geographic are of Victoria, Australia;
\item {}``\emph{in writing}'' includes communication via electronic means,
or by letter;
\item {}``\emph{electronic communication}'' means the same as in the Electronic Communications Act 2000; and
\item words importing one gender shall include all genders, and the singular
shall include the plural and \emph{vice versa}.
\end{itemize}

Unless the context otherwise requires, words or expressions contained
in these Articles have the same meaning as in the Act, but excluding
any statutory modification not in force when these Articles become
binding on the Association.

Apart from the exception mentioned in this Clause, any reference to an Act of Parliament includes any statutory modification or re-enactment of it for the time being in force.

\end{enumerate}

\end{document}