### 2014-11-17 Articles_of_Association_Melbourne_Bitcoin_Technology_Centre_MBTC_2015.tex
Add articles scaffold derrived from legislative instrument

% --------------
% LaTeX Document
% --------------

\documentclass[english,a4paper,titlepage,oneside]{article}

% --------
% PREAMBLE
% --------

% Use Packages
% ------------

% Language package
\usepackage[english]{babel}
% Input character encoding
\usepackage[utf8]{inputenc}
% Font encoding package
\usepackage{fontenc}
% Document layout package
\usepackage[top=2.25cm, bottom=3cm, left=2.5cm, right=2.5cm]{geometry}
% PDF document link package
\usepackage{hyperref}

% Document Settings
% -----------------

% Set paragraph vert space and indetatoin.
\setlength\parskip{\medskipamount}
\setlength\parindent{0pt}

% Assert paragraph unit formatting
\raggedright

% Hyphen dissuasion threshold
\hyphenpenalty=10000

% Define interperatation of @ charachter
\makeatletter
\makeatother

% ----------------
% DOCUMENT CONTENT
% ----------------

\begin{document}

% Cover page

\begin{titlepage}

  \author{BITCOIN TECHNOLOGY CENTERS, INC.} 
  \title{Articles Of Association} 
  \date{Mon 17 Nov, 2014} 
  \maketitle

  \section*{\centerline{Articles of Association of}}
  \section*{\centerline{BITCOIN TECHNOLOGY CENTERS, INC.}}
  \section*{\centerline{Incorporated in Victoria under the}}
  \section*{\centerline{Associations Incorporation Reform Act 2012}}

  The regulations as were in force upon the date of adoption of these Articles shall apply to the Association save to the extent that they are excluded by or are inconsistent with any of these Articles.

\end{titlepage}

% Cover page
% Index page
% Content pages

\section*{
  \begin{enumerate}
    \item Interpretation
    \linebreak[1]
    \textnormal{
    \\In these articles:
    \begin{enumerate}
      \item{} ``\emph{the Act}'' means the Associations Incorporation Reform Act 
      2012 including the Associations Incorporation Reform Regulations 2012;
      \item {}``\emph{address}'' means a postal address or, for the purposes
      of electronic communication, an e-mail address or an addressable 
      communication channel in each case registered with or by the Association;
      \item {}``\emph{the Association}'' means the association intended to be bound by these articles;
      \item {}``\emph{clear days}'' in relation to the period of a notice means
      a period excluding:
      \begin{enumerate}
        \item the day when the notice is given or deemed to be given; and
        \item the day for which it is given or on which it is to take effect;
      \end{enumerate}
      \item {}``\emph{the Articles}'' means the Articles of Association
      of the Association;
      \item {}``\emph{officers}'' includes the Board and the Secretary;
      \item {}``\emph{the seal}'' means the common seal of the Association if it
      has one;
      \item {}``\emph{Secretary}'' means the Secretary of the Association or any
      other person appointed to perform the duties of the Secretary of the
      Association, including a joint, assistant or deputy secretary;
      \item {}``\emph{the Board}'' means the members of the Association so appointed. During any period in which  the    Association is a charity, the Board members are trustees as defined by Section the Charities Act 1997;
      \item {}``\emph{Victoria}'' means the geographic are of Victoria, Australia;
      \item {}``\emph{in writing}'' includes communication via electronic means,
      or by letter;
      \item {}``\emph{electronic communication}'' means the same as in the Electronic Communications Act 2000; and
      \item words importing one gender shall include all genders, and the singular
      shall include the plural and \emph{vice versa}.
    \end{enumerate}
    } % end textnormal
    \item Identity
    \textnormal{
    \begin{enumerate}
      \item The name of the incorporated association, such as "XYZ Association Inc".
    \end{enumerate}
    }
    \item Purpose
    \textnormal{
    \begin{enumerate}
      \item The purposes of the incorporated association. These explain what the incorporated association is established to do.
    \end{enumerate}
    } % end textnormal
    \item Membership
    \textnormal{
    \begin{enumerate}
      \item The qualifications (if any) for membership of the incorporated association. If there are no specific qualifications for membership, the rules should say so. We recommend that your rules set out how someone applies to become a member; for example by making a request in writing to the secretary of the association.
    \end{enumerate}
    } % end textnormal
    \item Membership Fees
    \textnormal{
    \begin{enumerate}
      \item The entrance fees, subscriptions and other amounts (if any) to be paid by members of the incorporated association. If there are no membership fees, you should state that no fees, subscriptions or other payments are required from members. (Item 20 covers the association's source of funds. If no fees are charged, you should not mention fees in your rule addressing item 20.)
    \end{enumerate}
    } % end textnormal
    \item Membership Rights
    \textnormal{
    \begin{enumerate}
      \item The rights, obligations and liabilities of members. This must cover:
      \begin{enumerate}
        \item - which members have the right to vote at general meetings; there must be at least five voting members.
        \item- who will be informed of, and is permitted to attend, general meetings
        \item- who can use facilities of the association
        \item- financial obligations of members.
      \end{enumerate}
    \end{enumerate}
    } % end textnormal
    \item Membership Termination
    \textnormal{
    \begin{enumerate}
      \item Provisions for the resignation of a member or cessation of membership. For example, they might have to give written notice to the secretary or to a committee member. They may cease to be a member if their annual subscription is overdue for a substantial period, or they fail to attend a certain number of annual general meetings.
    \end{enumerate}
    } % end textnormal
    \item Dispute Resolution
    \textnormal{
    \begin{enumerate}
      \item The grievance procedures for settling disputes under the rules between the incorporated association and any members or between a member and other member. This procedure must give each party to the dispute an opportunity to be heard on the matter and ensure that an unbiased decision maker decides the outcome of the dispute.
    \end{enumerate}
    } % end textnormal
    \item Management Group
    \textnormal{
    \begin{enumerate}
      \item The name, membership and powers of the committee or other managing body.
    \end{enumerate}
    } % end textnormal
    \item Management Appointment
    \textnormal{
    \begin{enumerate}
      \item (a) Set out a process for the election or appointment of members of the committee (or governing board) of the association.
    \linebreak[1]
      \item The rules must identify the name by which the governing body is known (this may simply be ?the committee?) and the membership of the committee, such as president, vice president, treasurer or secretary. Such positions are commonly referred to as the office bearers of the association. In addition to office bearers, many associations also have ordinary members on the committee who do not hold a formal title. You may wish to have separate provisions to address the election or appointment of office bearers and ordinary members (if any) to the committee.
    \end{enumerate}
    } % end textnormal
    \item Management Terms
    \textnormal{
    \begin{enumerate}
      \item (b) The terms of office of members of the committee. For example, members may hold office for a term of one year and then be eligible for re-election or reappointment at the next annual general meeting.
      \item (c) Reasons why the office of a member of the committee becomes vacant. Generally these include the end of term, resignation, death, insolvency (bankruptcy) or loss of mental capacity or being removed from office by a resolution of the association at a general meeting.
      \linebreak[1]
      \item A member of the committee of an association vacates office in the circumstances provided in the rules of the association and if the member:
      \linebreak[1]
      \item - resigns by written notice addressed to the committee
      \item - is removed from office by a special resolution
      \item - dies
      \item - becomes insolvent
      \item - becomes a represented person under the Guardianship and Administration Act 1986 (for example, because they suffered an accident that caused a brain injury).
    \end{enumerate}
    } % end textnormal
    \item Interim Management
    \textnormal{
    \begin{enumerate}
      \item (d) The filling of casual vacancies occurring within the committee. This is where a member of the committee ceases to hold office before their term of appointment ends (such as if they resign or die). Your rules could include a process for the committee or president to appoint someone to fill the office until the term for that position ends.
    \end{enumerate}
    } % end textnormal
    \item Meetings
    \textnormal{
    \begin{enumerate}
      \item (e) The quorum and procedure at meetings of the committee. The quorum is the minimum number of committee members who must be present before a committee meeting can be held. You can express this as a whole number or as a percentage or fraction.
      \linebreak[1]
      \item Your rules must also cover the procedure to be followed at committee meetings, such as:
      \linebreak[1]
      \item - how much notice committee members will be given before meetings
      \item - who chairs committee meetings
      \item - order of business
      \item - how decisions are made. Do they need to be unanimous or is there a vote? Does the Chair have a casting vote?
      \item - what should happen if a committee member has a conflict of interest regarding a matter under consideration
      \item - can meetings be conducted using telephone or video communications or does every member have to be physically present in the same place
      \item - who takes minutes of the meeting and what is to be recorded.
    \end{enumerate}
    } % end textnormal
    \item Secretary Cessation
    \textnormal{
    \begin{enumerate}
      \item The procedures for the appointment and removal of the secretary. If the secretary is a member of the committee of your association under your rules, then the procedure for appointment and removal of committee members will be sufficient to cover this.
    \end{enumerate}
    } % end textnormal
    \item Records
    \textnormal{
    \begin{enumerate}
      \item The custody of records, securities and other relevant documents of the incorporated association. Set out who is responsible for this.
    \end{enumerate}
    } % end textnormal
    \item Common Seal
    \textnormal{
    \begin{enumerate}
      \item Provisions for the custody and use of the association’s common seal (if any). Use of a common seal is optional. If your association has one, the rules must state who is responsible for custody of the seal.
      \item Provisions for members to have access to, and to be able to obtain copies of, the records, securities and other relevant documents. These rules should cover:
      \linebreak[1]
      \item - information relating to incorporation, rules, management, membership records and financial statements
      \item - the association’s transactions, dealings, business or property
      \item - which records the committee may refuse to permit inspection of, such as confidential personal, employment, commercial or legal matters
      \item - the right to inspect the register of members, at a reasonable time.
      \linebreak[1]
      \item Under the Act, a member is entitled to inspect the rules of their association and minutes of general meetings of the association at any reasonable time. They are also entitled to a copy of the rules of their association or minutes of general meetings if they make a request in writing to their association for a copy.
      \linebreak[1]
      \item If a member requests to inspect the register of members, the incorporated association must allow this at a reasonable time.
      \linebreak[1]
      \item Your rules may also provide for a member to be able to obtain a copy of the register of members but this is optional – you could provide for inspection but not allow copying.
    \end{enumerate}
    } % end textnormal
    \item Minutes of Meetings
    \textnormal{
    \begin{enumerate}
      \item The preparation and retention (custody and storage) of accurate minutes of (see following two provisions):
      \linebreak[1]
      \begin{enumerate}
        \item (a) general meetings of the incorporated association.
        \linebreak[1]
        \item (b) meetings of the committee or other body having the management of the incorporated association.
      \end{enumerate}
    \end{enumerate}
    } % end textnormal
    \item Transparency \& Disclosure
    \textnormal{
    \begin{enumerate}
      \item Provision for members to have access to, and to be able to obtain copies of, minutes of general meetings, including financial statements that will be submitted at the association’s annual general meeting.
    \end{enumerate}
    } % end textnormal
    \item Administration
    \textnormal{
    \begin{enumerate}
      \item Right of access (if any) members have to minutes of meetings of the committee, including any terms and conditions subject to which access may be granted. You do not have to allow members access to minutes of committee meetings, but you must state whether or not you will allow it.
    \end{enumerate}
    } % end textnormal
    \item General Meetings
    \textnormal{
    \begin{enumerate}
      \item State how many general meetings the association must have each year and the period between those meetings. For example, the association may hold a general meeting every four months. The Act only requires an association to hold one general meeting per year, being the annual general meeting. Holding additional general meetings is decided by each association.
      \linebreak[1]
      \item Your rules must also provide for the manner in which a general meeting may be called. For example, general meetings may be called by the secretary or the committee or by a specified number or percentage of members giving a request in writing to the secretary to call a general meeting.
    \end{enumerate}
    } % end textnormal
    \item Conduct of Meetings
    \textnormal{
    \begin{enumerate}
      \item Set out the procedure to be followed at general meetings. Include:
      \linebreak[1]
      \item - who chairs the meeting
      \item - how voting is conducted (show of hands, secret ballot, casting vote)
      \item - whether and in what circumstances the meeting can be adjourned
      \item - the minimum number or percentage of all members who must be present to conduct a valid general meeting (the quorum)
      \item - whether voting by proxy is allowed.
    \end{enumerate}
    } % end textnormal
    \item Notice of Meetings
    \textnormal{
    \begin{enumerate}
      \item Set out the period of notice required to be given to members of a forthcoming general meeting and the manner in which notice is to be given (such as post or email or both). The rules must also set out the period of notice (if any) a member must give other members if they propose to move a motion at a general meeting. If you wish to allow for members to propose motions from the floor at a general meeting, state that no advance notice is required.
      \linebreak[1]
      \item The Act imposes specific requirements for notice where the motion requires a special resolution for it to be passed.
      \linebreak[1]
      \item A motion proposing an alteration to the rules of an incorporated association must be passed by a special resolution.
      \linebreak[1]
      \item A special resolution must be passed by at least 75\% of the members present or voting by proxy at a general meeting
      \linebreak[1]
      \item Members must be given at least 21 days’ notice of a motion that is to be passed by special resolution. The notice must include:
      \linebreak[1]
      \item - the date, time and place of the meeting
      \item - the full proposed resolution
      \item - a statement of the intention that the motion be proposed as a special resolution.
    \end{enumerate}
    } % end textnormal
    \item Financial Resources
    \textnormal{
    \begin{enumerate}
      \item The sources from which funds of the incorporated association will be or may be derived. These may include fees paid by members, grants and donations, proceeds from the sale of products or materials (if relevant).
    \end{enumerate}
    } % end textnormal
    \item Financial Management
    \textnormal{
    \begin{enumerate}
      \item The manner in which the funds of the incorporated association must be managed and, in particular, the mode of drawing and signing cheques on behalf of the incorporated association. Include who:
      \linebreak[1]
      \item - is responsible for receiving funds on behalf of the association and issuing receipts for those funds
      \item - is responsible for paying funds received into the associations’ bank account
      \item - can authorise expenditure by the association and how authorisation is given (such as by signature of the treasurer and another committee member)
      \item - can sign cheques on behalf of the association and what authorisation they need to do so (such as resolution of the committee).
    \end{enumerate}
    } % end textnormal
    \item Altering These Articles
    \textnormal{
    \begin{enumerate}
      \item Provide for the process to be followed to alter your rules, including adding new rules and removing old rules. This must be consistent with the requirements of the Act and a simple statement that the rules of the association may be altered by special resolution at a general meeting of the association is sufficient.
    \end{enumerate}
    } % end textnormal
    \item Winding Up and Dissolution
    \textnormal{
    \begin{enumerate}
      \item Set out what is to happen to any surplus assets of the association if it is wound up or dissolved. The distribution of surplus assets must not be contrary to the Act and generally surplus assets must not be distributed to any member or former member of the association.
    \end{enumerate}
    } % end textnormal
  \end{enumerate}
  }
\end{document}